---
title: "Lab2"
author: "Kevin Banh"
date: "5/24/2020"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Lab 2

## Task 1

```{r}
getwd()
```

## Task 2

```{r}
# Set dird to working directory
dird="C:\\Users\\kevin\\OneDrive\\Documents\\MATH4753\\Lab2"
# Create a function that reads the files
myread=function(csv){
  fl=paste(dird,csv,sep="/")
  read.table(fl,header=TRUE,sep=",")
}
epagas=myread("EPAGAS.csv")
head(epagas)
```

## Task 3

```{r}
mpg <- c(epagas$MPG)
z=(mpg-mean(mpg))/sd(mpg)
mean(z)
var(z)
# The first result, mean(z), is essentially 0 and var(z) = 1, thus the results remain true
```
```{r}
# Find z values corresponding to possible outliers
z[abs(z)>=2 & abs(z)<=3]
# Find elements of the vector mpg that are possible outliers
mpg[abs(z)>=2 & abs(z)<=3]
```

```{r}
library(lattice)
mycol <- ifelse(abs(z)>3, "Red",
        ifelse(abs(z)>=2 &abs(z)<=3,"Blue", "Black"))

dotplot(mpg, col=mycol, main="Miles per gallon", xlab="Miles",ylab="Gallon")
```

## Task 4

```{r}
boxplot(mpg,main="Miles Per Gallon", col = "black",
        xlab= "Miles",
        ylab= "Gallon",
        horizontal = TRUE,
        notch = TRUE)
```

At least 3/4 of mpg will lie within 2 standard deviations of the mean
```{r}
# Find proportion of mpg that lie within 2 standard deviations of the mean
length(mpg[abs(z)<2])/length(mpg)

```
96% of mpg lie within 2 standard deviations therefore Chebyshev's theorem holds true.

The empirical rule states that at least 95% of mpg will lie within 2 standard deviations of their mean. We can see that from [1], 96% > 95% thus the empirical rule holds as well. This is because the data is relatively symmetric and not highly skewed due to a low number of outliers.


